---
title: "Energy integration slides"
author: "Cameron Roach"
date: "16/03/2017"
output: ioslides_presentation
---


```{r setup, include=FALSE} 
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	dev = "png"
)

rm(list=ls())

require(dplyr)
require(tidyr)
require(lubridate)
require(stringr)
require(ggplot2)
theme_set(theme_bw())
require(readr)
require(quantreg)
require(foreach)
require(doMC)
require(knitr)
require(splines)

cache_file <- "./cache/slide_session.RData"
data_dir <- "./data/"

if (file.exists(cache_file)) {
  load(cache_file)
} else {
  # Register cores for parallel processing
  registerDoMC(3)
  
  # Load data
  b_uids <- c("BID1002", "BID1012")
  
  bldng_df <- NULL
  for (iB in b_uids) {
    bldng_df <- read_csv(paste0(data_dir, iB, ".csv"),
                         col_types = cols(
                           `LOCAL Time` = "c"
                         )) %>%
      rename(`Power (kW)` = `kW Power (kW)`) %>% 
      mutate(b_uid = iB,
             ts = dmy_hm(paste(Date, `LOCAL Time`)),
             date = date(ts),
             period = factor(hour(ts)*4 + minute(ts)/15 + 1),
             wday = factor(wday(ts, label = TRUE), ordered = FALSE),
             weekend = if_else(wday %in% c("Sat", "Sun"), TRUE, FALSE)) %>% 
      filter(!is.na(`Temperature (degrees Celsius)`)) %>% 
      # Add lagged summary statistics
      group_by(Date) %>% 
      mutate(mean_day_temp = mean(`Temperature (degrees Celsius)`,
                                  na.rm=TRUE)) %>% 
      ungroup() %>% 
      mutate(mean_day_temp_lag1 = lag(mean_day_temp, 96, order_by = ts),
             mean_day_temp_lag2 = lag(mean_day_temp, 192, order_by = ts)) %>% 
      select(-c(Date, `LOCAL Time`)) %>% 
      bind_rows(bldng_df)
  }
  
  
  # Fit working days only
  bldng_df <- bldng_df %>% 
    filter(weekend == FALSE)
  
  # Filter outliers (public holidays)
  # bldng_df %>%
  #   mutate(date = date(ts), year = year(ts)) %>%
  #   group_by(b_uid, year, date) %>%
  #   summarise(kW = mean(`Power (kW)`, na.rm=TRUE)) %>%
  #   ggplot(aes(x= factor(year), y=kW)) +
  #   geom_boxplot() +
  #   facet_wrap(~b_uid)
  
  outlier_days <- bldng_df %>% 
    mutate(date = date(ts), year = year(ts)) %>% 
    group_by(b_uid, year, date) %>% 
    summarise(kW = mean(`Power (kW)`, na.rm=TRUE)) %>% 
    group_by(b_uid, year) %>% 
    mutate(q1 = quantile(kW, 0.25),
           q2 = quantile(kW, 0.5),
           q3 = quantile(kW, 0.75),
           iqr = q3 - q1,
           outlier_low = q1 - 1.5 * iqr,
           outlier = if_else(kW < outlier_low, TRUE, FALSE)) %>% 
    ungroup()
  
  bldng_df <- outlier_days %>% 
    select(b_uid, date, outlier) %>% 
    inner_join(bldng_df) %>% 
    filter(outlier != TRUE)
  
  
  
  # Fit model
  #quantile_sequence <- 10:90/100
  quantile_sequence <- 1:9/10
  rq_fit <- foreach (iB = b_uids) %dopar% {
    rq(#`Power (kW)` ~ poly(`Temperature (degrees Celsius)`, 2) + `Humidity (%)` + period + mean_day_temp_lag1 + mean_day_temp_lag2,
      `Power (kW)` ~ ns(`Temperature (degrees Celsius)`, 2) + period + mean_day_temp_lag1 + mean_day_temp_lag2,
      tau = quantile_sequence,
      data = filter(bldng_df,
                    b_uid == iB))
  }
  names(rq_fit) <- b_uids
  
  # Predictions
  bldng_pred <- bldng_df %>% 
    group_by(b_uid) %>% 
    do(data.frame(predict(rq_fit[[.$b_uid[1]]], .),
                  .,
                  check.names = FALSE))
  
  dir.create("./cache", F, T)
  save.image(file = cache_file)
}
```


# Building Level Energy Forecasting

## Managing Commercial Buildings

Commercial buildings require energy forecasting to help:

- Manage peak demand.
- Quantify the impacts of building management changes.
- Assess performance and energy efficiency.

Buildings Alive works with 150+ commercial buildings which include supermarkets, hospitals and office blocks. Each require daily forecasts to inform facilities managers (FMs).

```{r ba-logo, fig.align="right", out.height="50%"}
include_graphics("./fig/ba-logo.pdf")
```

## Building Level Data

```{r energy-plot, fig.align="center", fig.cap="Energy usage for two commercial buildings."}
p_date_start <- dmy("16/1/2017")
p_date_end <- dmy("20/1/2017")

bldng_df %>% 
  filter(between(date(ts), p_date_start, p_date_end)) %>% 
  ggplot(aes(x=ts, y=`Power (kW)`)) + 
  geom_line() + 
  facet_wrap(~b_uid, ncol = 1) +
  xlab("Date")
```


## Building Level Data

```{r weather-energy-plot, fig.align="center", fig.cap="Natural cubic splines for each period of the day (df = 2)."}
bldng_df %>% 
  mutate(year = year(ts),
         period = as.numeric(period)) %>% 
  ggplot(aes(x = `Temperature (degrees Celsius)`, y = `Power (kW)`,
             colour = period, group = period)) +
  geom_smooth(method = "lm", formula = y ~ ns(x, df = 2), se = FALSE, alpha = 0.3) +
  facet_wrap(~b_uid) +
  scale_colour_gradient2(low = "red", mid = "springgreen3", high = "blue",
                         midpoint = 49) +
  labs(colour = "Period")
```



## Quantile Regression

Useful forecasts can be produced using quantile regression. Instead of only forecasting the conditional mean the entire distribution is forecast.

Uses the pinball loss function,

$$
L_\tau \left( y, q_\tau \right) =
\begin{cases}
  \tau(y-q_\tau) & \text{ for } y \geq q_\tau,\\
  (1-\tau)(q_\tau-y) & \text{ for } q_\tau > y.
\end{cases}
$$

```{r qr-ols-loss-functions, out.width="40%", fig.cap="Ordinary least squares and pinball loss functions.", fig.align="center", fig.show="hold"}
ggplot(data.frame(x=c(-1, 1)), aes(x)) +  
  stat_function(fun = function(x) {x^2}) + 
  #xlab("x") + 
  xlab("") +
  ylab("L") + 
  ggtitle("OLS loss function") 

ggplot(data.frame(x=c(-1, 1)), aes(x)) +  
  stat_function(xlim = c(-1, 0), 
                fun = function(x) {(1-0.8)*-x}) + 
  stat_function(xlim = c(0, 1), 
                fun = function(x) {0.8*x}) + 
  #xlab(expression(q[tau])) + 
  xlab("") +
  ylab(expression(L[tau] (y, q[tau]))) + 
  ggtitle("Pinball loss function") 
``` 

## Quantile Regression Forecasting

```{r quantile-plot, fig.align="center", fig.cap="Quantile forecasts for two commercial buildings."}
ggplot() +
  geom_line(data = bldng_pred %>% 
              filter(between(date(ts), p_date_start, p_date_end)) %>% 
              select(b_uid, ts, starts_with("tau")) %>% 
              gather(Quantile, `Power (kW)`, starts_with("tau")) %>% 
              mutate(Quantile = as.numeric(str_extract(Quantile, "\\d.\\d+"))),
            aes(x=ts, y=`Power (kW)`, colour = Quantile, group = Quantile),
            alpha = 0.5) +
  geom_line(data = bldng_df %>% 
              filter(between(date(ts), p_date_start, p_date_end)),
            aes(x=ts, y=`Power (kW)`), colour = "black") +
  facet_wrap(~b_uid, ncol = 1) +
  xlab("Date") +
  scale_colour_gradient(low = "springgreen3", high = "red")
```


## Assessing performance

* Forecasting a full distribution allows FMs to better assess risks and take appropriate actions.
* Allows FMs to know the severity __and probability__ of demand peaks.
* Can immediately assess if a building's performance was good compared to historical performance under similar conditions.